---
title: "Data Visualizations"
output: html_notebook
---
**EXERCISE 3.6.1**

**What geom would you use to draw a line chart? A boxplot? A histogram? An area chart?**
```{r}
library(tidyverse)
```
```{r}
ggplot(data = mpg)+
  geom_line(mapping=aes(x=displ,y=hwy))
  
```
```{r}
ggplot(data = mpg)+
  geom_boxplot(mapping=aes(x=displ,y=hwy,color=drv))
  
```
```{r}
ggplot(data = mpg)+
  geom_histogram()
  
```
```{r}
ggplot(data = mpg)+
  geom_area(mapping=aes(x=displ,y=hwy))
  
```
**Run this code in your head and predict what the output will look like. Then, run the code in R and check your predictions.**
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```
**What does show.legend = FALSE do?**
Removes a legend that explains which levels correspond to which values.
**What happens if you remove it?** 
You see the legend that explains which levels correspond to which values.
**Why do you think I used it earlier in the chapter?**
To show ho to remove legend
```{r}
ggplot(data = mpg) +
  geom_smooth(
    mapping = aes(x = displ, y = hwy, color = drv),
    show.legend = FALSE
  )
```
**What does the se argument to geom_smooth() do?**
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = TRUE)
```
**ANSWER**
se argument to geom_smooth()  does add a loess smooth
**Will these two graphs look different? Why/why not?**
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()
```
**ANSWER**
No the two graphs will not look different, because both functions produce the same output. The difference is that second function only does duplication in your code, which is not a good practice for obtaining clean code.
**Recreate the R code necessary to generate the following graphs.**
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(size=class),show.legend = FALSE) + 
  geom_smooth(se=FALSE)
```
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(size=class),show.legend = FALSE) + 
  geom_smooth(mapping = aes(group=drv),se=FALSE)
```
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color=drv,size=class),show.legend = TRUE) + 
  geom_smooth(mapping = aes(color=drv),show.legend = TRUE,se=FALSE)
```
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color=drv,size=class),show.legend = TRUE) + 
  geom_smooth(show.legend = FALSE,se=FALSE)
```
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color=drv,size=class),show.legend = TRUE) + 
  geom_smooth(mapping = aes(linetype=drv),show.legend = TRUE,se=FALSE)
```
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy,size=class)) + 
  geom_point(mapping = aes(color=drv),show.legend = TRUE) 
 
```